### 2020-09-16 Manuscript/manuscript.tex
add Kalahari paleogeographic neighbor subsection

\linenumbers

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../Figures/Fig1_map.pdf}
\caption{\small{\textbf{Simplified map of Laurentia showing the location of Archean provinces (labeled with text) and younger Paleoproterozoic and Mesoproterozoic crust (simplified from \citealp{Whitmeyer2007a} with additions for Greenland based on \citealp{St-Onge2009a}). Proterozoic orogens are labeled with \textit{italicized text} (or. ? orogen). The localities from which the compiled Precambrian paleomagnetic poles were developed are shown and colored by age. The circles (A rated poles) and squares (B rated poles) have been assessed by the Nordic workshop panel.}}}
\label{fig:Laurentia_map}
\end{figure}

@@

Following the Trans-Hudson orogeny, the locus of orogenesis migrated to the exterior of Laurentia. This change marks a shift in the predominant style of Laurentia's growth as subsequent crustal growth occurred dominantly through accretion of juvenile crust along the southern and eastern margin of the nucleus of Archean provinces (\citealp{Whitmeyer2007a}; Figs. \ref{fig:Laurentia_map} and \ref{fig:tectonic_history}). Determining the extent of these belts is complicated by poor exposure of them in the midcontinent relative to the exposure of the Archean provinces throughout the Canadian shield. Major growth of Laurentia following the amalgamation of these Archean provinces occurred associated with the arc-continent collision of the ca. 1.71 to 1.68 Ga Yavapai orogeny (Fig. \ref{fig:tectonic_history}). Yavapai orogenesis is interpreted to have resulted from the accretion of a series of arc terranes that collided with each other and Laurentia \citep{Karlstrom2001a}. Yavapai accretion was followed by widespread emplacement of granitoid intrusions \citep{Whitmeyer2007a}. These intrusions are hypothesized to have stabilized the juvenile accreted terranes that subsequently remained part of Laurentia \citep{Whitmeyer2007a}. Subsequent accretionary orogenesis of the ca. 1.65 to 1.60 Ga Mazatzal orogeny and associated plutonism led to further crustal growth in the latest Paleoproterozoic \citep{Karlstrom1988a}. Laurentia's growth continued in the Mesoproterozoic along the southeast margin through further juvenile terrane and arc accretion. An interval of major plutonism occurred ca. 1.48 to 1.35 Ga leading to the formation of A-type granitoids throughout both Mesoproterozoic and Paleoproterozoic provinces extending from the southwest United States up to the Central Gneiss Belt of Ontario to the northeast of Georgian Bay \citep{Slagstad2009a}. This plutonism is likely due to crustal melting within a back-arc region of ca. 1.50 to 1.43 Ga accretionary orogenesis \citep{Bickford2015a}. Geologic data from northern New Mexico have been interpreted to indicate an interval of ca. 1.49 to 1.40 Ga orogenesis that has been named the Picuris orogeny \citep{Daniel2013a}. Younger magmatic activity ca. 1.37 Ga of the Southern Granite-Rhyolite Province suggests a similar tectonic setting of accretionary orogenesis at that time \citep{Bickford2015a}. While an active margin interpretation with magmatism in a back-arc setting has gained traction within the literature, the tectonic setting is often described as enigmatic given earlier interpretations of an anorogenic setting (see references in \citealp{Slagstad2009a}).

@@

@@

@@

>>>>

%\linenumbers

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../Figures/Fig1_map.pdf}
\caption{\small{\textbf{Simplified map of Laurentia showing the location of Archean provinces (labeled with text) and younger Paleoproterozoic and Mesoproterozoic crust (simplified from \citealp{Whitmeyer2007a} with additions for Greenland based on \citealp{St-Onge2009a}). Proterozoic orogens are labeled with \textit{italicized text} (or. -- orogen). The localities from which the compiled Precambrian paleomagnetic poles were developed are shown and colored by age. The circles (A rated poles) and squares (B rated poles) have been assessed by the Nordic workshop panel.}}}
\label{fig:Laurentia_map}
\end{figure}

@@

Following the Trans-Hudson orogeny, the locus of orogenesis migrated to the exterior of Laurentia. This change marks a shift in the predominant style of Laurentia's growth as subsequent crustal growth occurred dominantly through accretion of juvenile crust along the southern and eastern margin of the nucleus of Archean provinces (\citealp{Whitmeyer2007a}; Figs. \ref{fig:Laurentia_map} and \ref{fig:tectonic_history}). Determining the extent of these belts is complicated by poor exposure of them in the midcontinent relative to the exposure of the Archean provinces throughout the Canadian shield. Major growth of Laurentia following the amalgamation of these Archean provinces occurred associated with the arc-continent collision of the ca. 1.71 to 1.68 Ga Yavapai orogeny (Fig. \ref{fig:tectonic_history}). Yavapai orogenesis is interpreted to have resulted from the accretion of a series of arc terranes that collided with each other and Laurentia \citep{Karlstrom2001a}. Yavapai accretion was followed by widespread emplacement of granitoid intrusions \citep{Whitmeyer2007a}. These intrusions are hypothesized to have stabilized the juvenile accreted terranes that subsequently remained part of Laurentia \citep{Whitmeyer2007a}. Subsequent accretionary orogenesis of the ca. 1.65 to 1.60 Ga Mazatzal orogeny and associated plutonism led to further crustal growth in the latest Paleoproterozoic \citep{Karlstrom1988a}. Laurentia's growth continued in the Mesoproterozoic along the southeast margin through further juvenile terrane and arc accretion. An interval of major plutonism occurred ca. 1.48 to 1.35 Ga leading to the formation of A-type granitoids throughout both Mesoproterozoic and Paleoproterozoic provinces extending from the southwest United States up to the Central Gneiss Belt of Ontario to the northeast of Georgian Bay \citep{Slagstad2009a}. This plutonism is interpreted to be associated with a combination of continental arc magmatism and melt generation within a back-arc region Laurentia's long-live active margin \citep{Bickford2015a}. Metamorphic rocks from Mesoproterozoic sedimentary and volcanic protoliths in northern New Mexico have been interpreted to indicate an interval of ca. 1.46 to 1.40 Ga orogenesis that has been named the Picuris orogeny \citep{Daniel2013a, Aronoff2016a}. Younger magmatic activity ca. 1.37 Ga of the Southern Granite-Rhyolite Province suggests a similar active margin setting at that time \citep{Bickford2015a}. While an active margin interpretation for the Granite-Rhyolite Province with arc and back-arc magmatism in a back-arc setting has gained traction within the literature, the tectonic setting is often described as enigmatic given earlier interpretations of an anorogenic setting (see references in \citealp{Slagstad2009a}).

@@

\subsubsection{Australia}

Could discuss Gibson2019b model related to hypothesized Paleoproterozoic tectonic history

@@

\subsubsection{Kalahari}

High-quality paleomagnetic poles constrain the coeval Umkondo large igneous province of the Kalahari craton and the early flood basalts of the Midcontinent Rift of Laurentia to have been separated by more than 50\textdegree$\;$ of latitude at the time they were emplaced with the craton margins separated by more than 30\textdegree$\;$ of latitude \citep{Swanson-Hysell2015a}. These data make it difficult to envision a shared origin of magmatism and pose a challenge to approaches that seek to reconstruct paleogeography on the basis on ages of LIPs alone. However, with the subsequent rapid motion of Laurentia to low latitudes it is possible that the Kalahari was a conjugate craton to the (south)east margin of Laurentia during the time of the Grenvillian orogeny. This conjugate relationship based on the interpretation that the Namaqua-Natal belt in southern Kalahari records late Mesoproterozoic collisional orogenesis was  proposed in \cite{Hoffman1991a} and is implemented in many reconstructions of Rodinia (e.g. \citealp{Li2008a}). Whether the Grenvillian margin of Laurentia and the Namaqua-Natal belt of Kalahari faced one another and been conjugates can be evaluated by paired paleomagnetic and geochronologic data sets from the Umkondo LIP and the Midcontinent Rift. The preferred interpretation of \cite{Swanson-Hysell2015a} and \citep{Kasbohm2015a} was that sites with northerly declinations from the Umkondo Province correspond to the reversed polarity directions from the early magmatic stage in the Midcontinent Rift (e.g. \citealp{Swanson-Hysell2014a}) such that Namaqua-Natal margin faced the Grenvillian margin of Laurentia. The late Mesoproterozoic apparent polar wander paths for Laurentia and Kalahari are consistent with them become conjoined as in Figure \ref{fig:Grenville_reconstructions} \citep{Swanson-Hysell2015a}. The record of the Namaqua belt wherein there is granitoid plutonism and arc accretion up to ca. 1090 followed by peak granulite metamorphism ca. 1065 to 1045 Ma \citep{Diener2013, Spencer2015a} is consistent with a scenario wherein Kalahari was on the upper plate that collided with Laurentia at the time of the Ottawan phase of Grenvillian orogenesis following subduction of oceanic lithosphere associated with an intervening ocean. If they indeed became conjoined in Rodinia, the separation of Kalahari from Laurentia likely initiated ca. 795 Ma heralded by the emplacement of the Gannakouriep diabase dike swarm \citep{Rioux2010a, DeKock2021a}.

@@

\noindent\textbf{large igneous province (LIP) } Voluminous and rapidly emplaced volcanics and intrusions that are typically of mafic composition. These provinces are often interpreted to be due to decompression melting of an upwelling mantle plume.
